\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-10-12 redaction/dhp_july2017.tex
changement paragraphe lashkaripour

Following \citet{samuelson1954}, standard models of international trade have usually modelled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, the structure (additive vs ad-valorem) of transport costs is far from being anecdotal. The literature has long emphasized its role in shaping the pattern of trade flows. As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. These findings deliver strong empirical support in favor of the Alchian-Allen conjecture. In contrast, highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result. The positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}, \citet{Lashkaripour-mimeo-2017} finds, based on firm-transaction-level data on Colombian imports, an average transport cost elasticity of 0.9, supporting the \emph{ad-valorem} assumption. The result if further confirmed on more aggregate US, sector-level trade data for the period 1989-1994. However, this contradicts several recent empirical studies based on micro-level data, supporting the existence of additive costs (i.e., cost per unit exported). Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis. Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs are significant, representing 14\% of the median price. Based on US sectoral similar to the one used by \citet{Lashkaripour-mimeo-2017}, but for a longer a period of time, our own estimates suggest also a more important part for additive costs.

>>>>

Following \citet{samuelson1954}, standard models of international trade have usually modelled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, the structure (additive vs ad-valorem) of transport costs is far from being anecdotal. The literature has long emphasized its role in shaping the pattern of trade flows. As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. \citet{Lashkaripour-mimeo-2017} challenges this view. Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result. While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports. This result is also confirmed on more aggregate US, sector-level trade data. The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis. Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs represent a sizeable share of international trade costs. Based on US sectoral data, our own estimates also suggest an important part for additive costs in international transport costs.
